---
title: "Museum Data Join"
name: "Holland Stam"
date: "`r Sys.Date()`"
---

```{r}
#| label: load-pkg
#| message: false

library(tidyverse)
library(tidymodels)
library(knitr)
library(janitor) # for the clean_names function
library(reshape)
library(ggplot2)
library(dplyr)

```

# Data prep

```{r}
#| label: load-data
#| message: false

# load gardner
gardner <- read_csv(here::here("Chapter1", "data/gardner.csv")) %>%
  janitor::clean_names()

# load janson
janson <- read_csv(here::here("Chapter1", "data/janson.csv")) %>%
  janitor::clean_names()

# load MoMA
moma <- read_csv(here::here("Chapter1", "data/moma.csv")) %>%
  janitor::clean_names() %>%
  filter(year != "MISSING YEAR") %>% 
  mutate(year = as.numeric(year))

# load CountMoMAExh
#countmoma <- read_csv(here::here("Chapter1", "data/countmoma.csv")) %>%
#  janitor::clean_names()

# load Whitney
whitney <- read_csv(here::here("Chapter1", "data/whitney.csv")) %>%
  janitor::clean_names()
```

**NOTE: Moma data has some (very few) rows where the year is `MISSING YEAR`, check if those can be filled in.**

Join `gardner` and `janson` with column = book, janson, gardner

```{r janson}
janson <- janson %>%
  mutate(
    artist_nationality_other =
      case_when(
        artist_nationality == "French" ~ "French",
        artist_nationality == "British" ~ "British",
        artist_nationality == "American" ~ "American",
        artist_nationality == "Spanish" ~ "Spanish",
        artist_nationality == "German" ~ "German",
        TRUE ~ "Other"
      )
  ) %>%
  select(
    "artist_name",
    "edition_number",
    "year",
    "artist_unique_id",
    "artist_nationality",
    "artist_nationality_other",
    "artist_gender",
    "artist_race",
    "artist_ethnicity",
    "book",
    "space_ratio_per_page"
  )
```

```{r gardner}
gardner <- gardner %>%
  mutate(
    artist_nationality_other =
      case_when(
        artist_nationality == "French" ~ "French",
        artist_nationality == "British" ~ "British",
        artist_nationality == "American" ~ "American",
        artist_nationality == "Spanish" ~ "Spanish",
        artist_nationality == "German" ~ "German",
        TRUE ~ "Other"
      )
  ) %>%
  select(
    "artist_name",
    "edition_number",
    "year",
    "artist_unique_id",
    "artist_nationality",
    "artist_nationality_other",
    "artist_gender",
    "artist_race",
    "artist_ethnicity",
    "book",
    "space_ratio_per_page"
  )
```

**NOTE: I don't think we want these joined, we want them to be binded to each other. So I changed that.**

```{r gardnerjanson}
gardnerjanson <- gardner %>%
  bind_rows(janson)

sumgj <- gardnerjanson %>%
  group_by(artist_name, year)%>%
  summarize(space_ratio_per_page_total = sum(space_ratio_per_page))

gardnerjanson <- gardnerjanson%>%
  left_join(sumgj)%>%
  select(-space_ratio_per_page)%>%
  unique()

```

------------------------------------------------------------------------

Create moma with complete years

```{r momaexh}
moma_complete_years <- moma %>%
  count(artist_name, year)%>%
  mutate(count = 1)%>%
  arrange(artist_name, year)%>%
  group_by(artist_name)%>%
  ungroup() %>%
  complete(artist_name, year = c(min(year):max(year))) %>%
  mutate(
    moma_count = if_else(is.na(count), 0, count)
    ) %>%
  arrange(artist_name, year)%>%
  group_by(artist_name)%>%
  mutate(moma_count_to_date = cumsum(moma_count))%>%
  ungroup()%>%
  select(artist_name, year, moma_count, moma_count_to_date)
```

Create whitney with complete years

```{r whitneyexh}
whitney_complete_years <- moma %>%
  count(artist_name, year)%>%
  mutate(count = 1)%>%
  arrange(artist_name, year)%>%
  group_by(artist_name)%>%
  ungroup() %>%
  complete(artist_name, year = c(min(year):max(year))) %>%
  mutate(
    whitney_count = if_else(is.na(count), 0, count)
    ) %>%
  arrange(artist_name, year)%>%
  group_by(artist_name)%>%
  mutate(whitney_count_to_date = cumsum(whitney_count))%>%
  ungroup()%>%
  select(artist_name, year, whitney_count, whitney_count_to_date)
```



Join moma and whitney info to gardnerjanson

**NOTE: I would expect gardnerjanson_museums to have the same number of rows as gardnerjanson, we should look into why that's not happening.**

This is occurring as there are some examples of an artists count of exhibitions per year

```{r museum-join}
gardnerjanson_museums <- gardnerjanson %>%
  left_join(moma_complete_years, by = c("artist_name", "year")) %>%
  left_join(whitney_complete_years, by = c("artist_name", "year"))
```

Make clean Artist ID - alphabetical for all artists across books

```{r artist-id}
#The artist_unique_id is alphabetrical per book - I should standardize it so that it is alphabetical regardless of book 

```

```{r exhcatcolumnmoma}
gardnerjanson_museums <- gardnerjanson_museums %>%
  mutate(moma_cat = case_when(moma_count_to_date >= 1 ~ '1', 
         TRUE ~ '0'))
```

```{r exhcatcolumnwhitney}
gardnerjanson_museums <- gardnerjanson_museums %>%
  mutate(whitney_cat = case_when(whitney_count_to_date >= 1 ~ '1', 
         TRUE ~ '0'))

```
# Exploratory data analysis

## Bivariate Plots and MLR with space ratio

Janson:

```{r countjanson}
ggplot(gardnerjanson_museums %>% filter(book == "janson"), 
       aes(x = space_ratio_per_page_total, fill = edition_number)) +
  geom_histogram() +
  labs(
    title = "Distribution of Ratio of Space\nper Artist per Edition per Page in Janson",
    x = "Ratio of Space per Page",
    y = "Count"
  )

summary(janson$space_ratio_per_page)

median <- median(janson$space_ratio_per_page)
```

The visualization above is right skewed, unimodal and assymetrical. Therefore, we would want to look at the median and IQR. The median of `space_ratio_per_page_total` in Janson is `r median`.

space_ratio_per_page with Gender

```{r genderjanson}
ggplot(gardnerjanson_museums %>% filter(book == "janson"), 
       aes(x = artist_gender, y = space_ratio_per_page_total)) +
  geom_boxplot() +
  labs(
    title = "Ratio of Space per Page per Artist per Edition\nin Janson by Gender",
    x = "Gender",
    y = "Ratio of Space per Page"
  )
```

space_ratio_per_page with Race

```{r racejanson}
ggplot(gardnerjanson_museums %>% filter(book == "janson"), 
       aes(x = artist_race, y = space_ratio_per_page_total)) +
  geom_boxplot() +
  labs(
    title = "Ratio of Space per Page per Artist per Edition\nin Janson by Race",
    x = "Race",
    y = "Ratio of Space per Page"
  )
```

space_ratio_per_page with Ethnicity

```{r jansonethnicity}
ggplot(gardnerjanson_museums %>% filter(book == "janson"), 
       aes(x = artist_ethnicity, y = space_ratio_per_page_total)) +
  geom_boxplot() +
  labs(
    title = "Ratio of Space per Page per Artist per Edition\nin Janson by Ethnicity",
    x = "Ethnicity",
    y = "Ratio of Space per Page"
  )
```

space_ratio_per_page with Nationality

```{r jansonnationality}
ggplot(gardnerjanson_museums %>% filter(book == "janson"), 
       aes(x = artist_nationality_other, y = space_ratio_per_page_total)) +
  geom_boxplot() +
  labs(
    title = "Ratio of Space per Page per Artist per Edition\nin Janson by Nationality",
    x = "Nationality",
    y = "Ratio of Space per Page"
  )
```

MLR of Janson with Gender +Race + Nationality + Ethnicity \~ space_ratio_per_page

```{r jansonMLR}
jansonMLR <- lm(space_ratio_per_page_total ~ artist_race +
  artist_ethnicity +
  artist_gender +
  artist_nationality_other +
  moma_count_to_date +
  whitney_count_to_date,
  data = gardnerjanson_museums %>% filter(book == "janson")
)
summary(jansonMLR)
```

```{r jansonrsquared}
glance(jansonMLR)$r.squared
```

```{r jansonadjrsquared}
glance(jansonMLR)$adj.r.squared

```

Gardner:

```{r countgardner}
ggplot(
  gardnerjanson_museums %>% filter(book == "gardner"),
  aes(x = space_ratio_per_page_total)
) +
  geom_histogram() +
  labs(
    title = "Distribution of Ratio of Space\nper Artist per Edition per Page in Gardner",
    x = "Ratio of Space per Page",
    y = "Count"
  )

gardnerjanson_museums %>%
  filter(book == "gardner") %>%
  summarize(
    min = min(space_ratio_per_page_total),
    q1 = quantile(space_ratio_per_page_total, 0.25),
    median = median(space_ratio_per_page_total),
    mean = mean(space_ratio_per_page_total),
    q3 = quantile(space_ratio_per_page_total, 0.75),
    max = max(space_ratio_per_page_total)
  )
```

space_ratio_per_page vs Gender

```{r gardnergender}
ggplot(gardnerjanson_museums %>% filter(book == "gardner"),
  aes(x = artist_gender, y = space_ratio_per_page_total)
) +
  geom_boxplot() +
  labs(
    title = "Ratio of Space per Page per Artist per Edition\nin Gardner by Gender",
    x = "Gender",
    y = "Ratio of Space per Page"
  )
```

space_ratio_per_page vs Race

```{r gardnerrace}
ggplot(
  gardnerjanson_museums %>% filter(book == "gardner"),
  aes(x = artist_race, y = space_ratio_per_page_total)
) +
  geom_boxplot() +
  labs(
    title = "Ratio of Space per Page per Artist per Edition\nin Gardner by Race",
    x = "Race",
    y = "Ratio of Space per Page"
  ) +
  theme(axis.text.x = element_text(angle = 45, size = 5)) +
  ylim(0, 2.0)
```

space_ratio_per_page vs Ethnicity

```{r gardnerethnicity}
ggplot(
  gardnerjanson_museums %>% filter(book == "gardner"),
  aes(x = artist_ethnicity, y = space_ratio_per_page_total)
) +
  geom_boxplot() +
  labs(
    title = "Ratio of Space per Page per Artist per Edition\nin Gardner by Ethnicity",
    x = "Ethnicity",
    y = "Ratio of Space per Page"
  )
```

space_ratio_per_page vs Nationality

```{r gardnernationality}
ggplot(
  gardnerjanson_museums %>% filter(book == "gardner"),
  aes(x = artist_nationality_other, y = space_ratio_per_page_total)
) +
  geom_boxplot() +
  labs(
    title = "Ratio of Space per Page per Artist per Edition\nin Gardner by Nationality",
    x = "Nationality",
    y = "Ratio of Space per Page"
  )
```

GardnerMLR Demographic

```{r gardnerMLR}
gardnerMLR <- lm(space_ratio_per_page_total ~ artist_race +
  artist_ethnicity +
  artist_gender +
  artist_nationality_other +
  moma_count_to_date +
  whitney_count_to_date,
data = gardnerjanson_museums %>% filter(book == "gardner")
)
summary(gardnerMLR)
```

```{r gardnerrsquared}
glance(gardnerMLR)$r.squared
```

```{r gardneradjrsquared}
glance(gardnerMLR)$adj.r.squared

```

MLR demograpic with jansongardner

```{r gardnerjansonMLR}
gardnerjansonMLR <- lm(space_ratio_per_page_total ~ artist_race +
  artist_ethnicity +
  artist_gender +
  artist_nationality_other +
  moma_count_to_date +
  whitney_count_to_date,
data = gardnerjanson_museums
)
summary(gardnerjansonMLR)
```

```{r gardnerjansonrsquared}
glance(gardnerjansonMLR)$r.squared
```

## Museums

**NOTE: I think something is wrong with the museum join, take a look to see if you can debug, or we can look together. Very few records are being matched, which seems odd.**

whitney eda

```{r whitneycounttodate}
ggplot(gardnerjanson_museums, aes(x = whitney_count_to_date)) +
  geom_histogram() +
  labs(
    title = "Distribution of Whitney Exhibitions per Artist per Edition",
    x = "Count to Date",
    y = "Count"
  )
```

```{r whitneyvisual}
ggplot(gardnerjanson_museums, aes(x = whitney_count_to_date, y = space_ratio_per_page_total)) +
  geom_point(position=position_jitter(h=0.1, w=0.1), alpha = 0.5, size = 1) +
  geom_smooth(method = "lm", se = FALSE)+
  labs(
    title = "Ratio of Space Per Artist Per Edition by\nNumber of Exhibitions per Artist per Editions in the Whitney",
    x = "Count to Date",
    y = "Ratio of Space Per Artist Per Edition"
  )

#corr_coef
```

```{r whitneymomaheatmap}
p <- ggplot(gardnerjanson_museums, aes(space_ratio_per_page_total,year)) + 
    geom_tile(aes(fill = whitney_count_to_date), colour = "steelblue") + 
    scale_fill_gradient(low = "steelblue",high = "red")
p
g <- ggplot(gardnerjanson_museums, aes(space_ratio_per_page_total,year)) + 
    geom_tile(aes(fill = moma_count_to_date), colour = "steelblue") + 
    scale_fill_gradient(low = "steelblue",high = "red")
g
```

moma eda

```{r momacounttodate}
ggplot(gardnerjanson_museums, aes(x = moma_count_to_date)) +
  geom_histogram() +
  labs(
    title = "Distribution of MoMA Exhibitions per Artist per Edition",
    x = "Count to Date",
    y = "Count"
  )
```

```{r momavisual}
ggplot(gardnerjanson_museums, aes(x = moma_count_to_date, y = space_ratio_per_page_total)) +
  geom_point(position=position_jitter(h=0.1, w=0.1), alpha = 0.5, size = 1) +
  geom_smooth(method = "lm", se = FALSE)+
  labs(
    title = "Ratio of Space Per Artist Per Edition by\nNumber of Exhibitions per Artist per Editions in the MoMA",
    x = "Count to Date",
    y = "Ratio of Space Per Artist Per Edition"
  )

#corr_coef
```
