---
title: "Museum Data Join"
name: "Holland Stam"
date: "02.01.2022"
---

```{r load-pkg-data}
#| message: false
library(tidyverse)
library(tidymodels)
library(knitr)

gardner <- read_csv("data/gardner.csv")
#loadjanson
janson <- read_csv("data/janson.csv")
#loadMoMA
moma <- read_csv("data/moma.csv")
#loadCountMoMAExh
countmoma <- read_csv("data/countmoma.csv")
#loadWhitney
whitney <- read_csv("data/whitney.csv")
```

join gardner and janson with column = book, janson, gardner

```{r janson}
janson <- janson %>%
    mutate(ArtistNationalityOther = 
             case_when((ArtistNationality != "French") &
                       (ArtistNationality != "American") &
                       (ArtistNationality != "German") &
                       (ArtistNationality != "Spanish") &
                       (ArtistNationality != "British") 
                                          ~ "Other",
             (ArtistNationality == "French" ~ "French"),
             (ArtistNationality == "British" ~ "British"),
             (ArtistNationality == "American" ~ "American"),
             (ArtistNationality == "Spanish" ~ "Spanish"),
             (ArtistNationality == "German" ~ "German"))) %>%
  select("ArtistName", 
        "EditionNumber",
        "Year", 
        "ArtistUniqueID", 
        "ArtistNationality",
        "ArtistNationalityOther",
        "ArtistGender", 
        "ArtistRace", 
        "ArtistEthnicity", 
        "Book", 
        "SpaceRatioPerPage")
```

```{r gardner}
gardner <- gardner %>%
    mutate(ArtistNationalityOther = 
             case_when((ArtistNationality != "French") &
                       (ArtistNationality != "American") &
                       (ArtistNationality != "German") &
                       (ArtistNationality != "Spanish") &
                       (ArtistNationality != "British") 
                                          ~ "Other",
             (ArtistNationality == "French" ~ "French"),
             (ArtistNationality == "British" ~ "British"),
             (ArtistNationality == "American" ~ "American"),
             (ArtistNationality == "Spanish" ~ "Spanish"),
             (ArtistNationality == "German" ~ "German"))) %>%
  select("ArtistName", 
        "EditionNumber",
        "Year", 
        "ArtistUniqueID", 
        "ArtistNationality",
        "ArtistNationalityOther",
        "ArtistGender", 
        "ArtistRace", 
        "ArtistEthnicity", 
        "Book", 
        "SpaceRatioPerPage")
```

```{r gardnerjanson}
gardnerjanson <- gardner %>%
  full_join(janson, by = c("ArtistName", 
                           "EditionNumber",
                           "Year", 
                           "ArtistUniqueID", 
                           "ArtistNationality", 
                           "ArtistNationalityOther",
                           "ArtistGender", 
                           "ArtistRace", 
                           "ArtistEthnicity", 
                           "Book", 
                           "SpaceRatioPerPage")) %>%
  select("ArtistName", 
        "EditionNumber",
        "Year", 
        "ArtistUniqueID", 
        "ArtistNationality",
        "ArtistNationalityOther",
        "ArtistGender", 
        "ArtistRace", 
        "ArtistEthnicity", 
        "Book", 
        "SpaceRatioPerPage")

```

countmomacleanNA

```{r countmomaclean}
countmoma <- countmoma %>%
  filter(`TotalAreaText(cm^2)`!= "NA")

```

join gardnerjanson with MoMA through antijoin - showing gardnerjanson that is NOT in MoMA - compare directly with moma?

```{r momaantijoin}
antimomagardnerjanson <- gardnerjanson %>%
  anti_join(moma)

```

clean artist name manually using gardnerjanson name

```{r whitneyantijoin}
antiwhitneygardnerjanson <- gardnerjanson %>%
  anti_join(whitney)

```

join gardnerjansonMoMa with Whitney through antijoin

arrange(ArtistName, EndYear)%>%

group_by(ArtistName)%>%

mutate(count_to_date = cumsum(count))%>%

select(ArtistName, count_to_date)

```{r momaexh}
moma <- moma %>%
  mutate(count = 1)%>%
  arrange(ArtistName, Year)%>%
  group_by(ArtistName)
```

```{r momagarderjanson}
momagardnerjanson <- gardnerjanson %>%
  left_join(moma, by = "ArtistName", "Year")

```

```{r whitneyexh}
whitney <- whitney %>%
  mutate(count = 1)%>%
  arrange(ArtistName, Year)%>%
  group_by(ArtistName)%>%
  mutate(count_to_date = cumsum(count))%>%
  select(ArtistName, count_to_date)
```

```{r whitneygardnerjanson}
whitneygardnerjanson <- gardnerjanson %>%
  right_join(whitney)

```

EDA: Bivariate Plots and MLR with SpaceRatio

Janson:

```{r countjanson}
ggplot(janson, aes(x = SpaceRatioPerPage, fill = EditionNumber))+
  geom_histogram()+
  labs(title = "Distribution of Ratio of Space 
       per Artist per Edition per Page in Janson",
       x = "Ratio of Space per Page",
       y = "Count")

summary(janson$SpaceRatioPerPage)

median(janson$SpaceRatioPerPage)

IQR(janson$SpaceRatioPerPage)

janson%>%
  arrange(desc(SpaceRatioPerPage))%>%
  select(ArtistName, EditionNumber, Year, SpaceRatioPerPage)%>%
  filter(SpaceRatioPerPage >= 1)



```

The visualization above is right skewed, unimodal and assymetrical. Therefore, we would want to look at the median to understand its center and IQR to understand its spread. The median of `SpaceRatioPerPage` in Janson is `r round(median(janson$SpaceRatioPerPage), 3)` and the IQR is `r round(IQR(janson$SpaceRatioPerPage), 3)`. There are potential outliers that are over 1.0, meaning the amount of space, area of text and area of figure in text combined is more than just the area of a single page within the respective book of *Janson's History of Art.*

```{r genderjanson}
ggplot(janson, aes(x=ArtistGender, y= SpaceRatioPerPage))+
  geom_boxplot()+
  labs(title = "Ratio of Space per Page per Artist per Edition in 
       Janson by Gender",
       x = "Gender",
       y = "Ratio of Space per Page")

janson_female <- janson %>%
  filter(ArtistGender == "Female")

median(janson_female$SpaceRatioPerPage)

janson_male <- janson %>%
  filter(ArtistGender == "Male")

median(janson_male$SpaceRatioPerPage)

```

Interestingly enough, the median `SpaceRatioPerPage` for female artists is `r round(median(janson_female$SpaceRatioPerPage), 3)`, not far below the median `SpaceRatioPerPage` for male artists, `r round(median(janson_male$SpaceRatioPerPage), 3)`in *Janson's History of Art.* This indicates that even though the number of female artists included is far fewer, `r nrow(janson_female)`, the median amount of space given to women is similar to that as the median amount of space given to male artists, `r nrow(janson_male)`. The IQR is larger for male artists than it is for female artists, and there are

SpaceRatioPerPage with Race

```{r racejanson}
ggplot(janson, aes(x=ArtistRace, y= SpaceRatioPerPage))+
  geom_boxplot()+
  labs(title = "Ratio of Space per Page per Artist per Edition in 
       Janson by Race",
       x = "Race",
       y = "Ratio of Space per Page")

```

SpaceRatioPerPage with Ethnicity

```{r jansonethnicity}
ggplot(janson, aes(x=ArtistEthnicity, y= SpaceRatioPerPage))+
  geom_boxplot()+
  labs(title = "Ratio of Space per Page per Artist per Edition in
       Janson by Ethnicity",
       x = "Ethnicity",
       y = "Ratio of Space per Page")
```

SpaceRatioPerPage with Nationality

```{r jansonnationality}
ggplot(janson, aes(x=ArtistNationalityOther, y= SpaceRatioPerPage))+
  geom_boxplot()+
  labs(title = "Ratio of Space per Page per Artist per Edition in 
       Janson by Nationality",
       x = "Nationality",
       y = "Ratio of Space per Page")
```

MLR of Janson with Gender +Race + Nationality + Ethnicity \~ SpaceRatioPerPage

```{r jansonMLR}
jansonMLR = lm(SpaceRatioPerPage ~ ArtistRace + 
                         ArtistEthnicity + 
                         ArtistGender +
                         ArtistNationalityOther, 
                       data = janson)
summary(jansonMLR)

```

```{r jansonrsquared}
glance(jansonMLR)$r.squared

```

Gardner:

```{r countgardner}
ggplot(gardner, aes(x = SpaceRatioPerPage))+
  geom_histogram()+
  labs(title = "Distribution of Ratio of Space 
       per Artist per Edition per Page in Gardner",
       x = "Ratio of Space per Page",
       y = "Count")

summary(gardner$SpaceRatioPerPage)
```

SpaceRatioPerPage vs Gender

```{r gardnergender}
ggplot(gardner, aes(x=ArtistGender, y= SpaceRatioPerPage))+
  geom_boxplot()+
  labs(title = "Ratio of Space per Page per Artist per Edition in 
       Gardner by Gender",
       x = "Gender",
       y = "Ratio of Space per Page")

```

SpaceRatioPerPage vs Race

```{r gardnerrace}
ggplot(gardner, aes(x=ArtistRace, y= SpaceRatioPerPage))+
  geom_boxplot()+
  labs(title = "Ratio of Space per Page per Artist per Edition in 
       Gardner by Race",
       x = "Race",
       y = "Ratio of Space per Page")+
   theme(axis.text.x = element_text(angle = 45, size = 5))+
  ylim(0, 2.0)
```

SpaceRatioPerPage vs Ethnicity

```{r gardnerethnicity}
ggplot(gardner, aes(x=ArtistEthnicity, y= SpaceRatioPerPage))+
  geom_boxplot()+
  labs(title = "Ratio of Space per Page per Artist per Edition in 
       Gardner by Ethnicity",
       x = "Ethnicity",
       y = "Ratio of Space per Page")
```

SpaceRatioPerPage vs Nationality

```{r gardnernationality}
ggplot(gardner, aes(x=ArtistNationalityOther, y= SpaceRatioPerPage))+
  geom_boxplot()+
  labs(title = "Ratio of Space per Page per Artist per Edition in 
       Gardner by Nationality",
       x = "Nationality",
       y = "Ratio of Space per Page")

```

GardnerMLR Demographic

```{r gardnerMLR}
gardnerMLR = lm(SpaceRatioPerPage ~ ArtistRace + 
                         ArtistEthnicity + 
                         ArtistGender +
                         ArtistNationalityOther, 
                       data = gardner)
summary(gardnerMLR)

```

```{r gardnerrsquared}
glance(gardnerMLR)$r.squared

```

MLR demograpic with jansongardner

```{r gardnerjansonMLR}
gardnerjansonMLR = lm(SpaceRatioPerPage ~ ArtistRace + 
                         ArtistEthnicity + 
                         ArtistGender +
                         ArtistNationalityOther, 
                       data = gardnerjanson)
summary(gardnerjansonMLR)
```

```{r gardnerjansonrsquared}
glance(gardnerjansonMLR)$r.squared

```

whitney eda

{r whitneycounttodate} ggplot(whitneygardnerjanson, aes(x = count_to_date))+ geom_histogram()+ labs(title = "Distribution of Whitney Exhibitions per Artist per Edition", x = "Count to Date", y = "Count")

{r whitneyvisual} ggplot(whitneygardnerjanson, aes(x = count_to_date, y = SpaceRatioPerPage))+ geom_point()+ labs(title = "Ratio of Space Per Artist Per Edition by Number of Exhibitions per Artist per Editions in the Whitney", x = "Count to Date", y = "Ratio of Space Per Artist Per Edition")

moma eda

{r momacounttodate} ggplot(momagardnerjanson, aes(x = count_to_date))+ geom_histogram()+ labs(title = "Distribution of MoMA Exhibitions per Artist per Edition", x = "Count to Date", y = "Count")

{r momavisual} ggplot(momagardnerjanson, aes(x = count_to_date, y = SpaceRatioPerPage))+ geom_point()+ labs(title = "Ratio of Space Per Artist Per Edition by Number of Exhibitions per Artist per Editions in the MoMA", x = "Count to Date", y = "Ratio of Space Per Artist Per Edition")
